### 2020-03-30 source/sec_convexsmoothsgd.tex
partial sgd v1

% !TEX ROOT=./main.tex

\textbf{Full Device Participation}

\begin{itemize}
	\item Stochastic gradient of device $k$ at time step $t$, at point $w_{t,k}$: 
	$$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$$
	$$ \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $$
	$$\EE \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}\right)$$
\item  One-step stochastic subgradient of all devices.
$$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $$
\begin{align}
	\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}
	\label{eq:egtsgd}
\end{align}
\end{itemize}

$$\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2 = \EE \|\overline{\vv}_{t+1} - \vw^*\|^2,$$
According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand $\Delta_{t+1}$
$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$

Take the expectation condition on $\vw_t$ over random samples at all devices:
\begin{align}
\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\ov{w}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
\label{eq:expandsgd}
\end{align}

Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:

\begin{align*}
	& -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*}\right> \\
 =  & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> -2 \eta_{t} \left<\EE \vg_{t}, \vw^{k}_t - \vw^{*}\right>\\
 \leq & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\\
 \leq &\sum_{k=1}^N p_k \left[2 \eta_{t} (F_k(\vw^{k}_t) - F_k(\ov{w}_t) + \frac{L}{2} \|\ov{w}_t - \vw^{k}_t\|^2 ) + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\right]\\
 = & \sum_{k=1}^N p_k \eta_t L \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} \sum_{k=1}^N p_k (F_k(w^*) - F_k(\ov{w}_t))\\
 = &  \eta_t L \sum_{k=1}^N p_k \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))
\end{align*}
Plug in this upper bound into \eq{\ref{eq:expandsgd}}, we have

\begin{align}
\EE\left\|\ov{w}_{t+1}-\vw^{*}\right\|^2 &\leq \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \eta_t L \sum_{k=1}^N p_k \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))\\
&\leq \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))\\
&\leq \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} G^2 +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - F(\ov{w}_t)) \label{eq:cvxsgd1}
\end{align}

Sum two sides of \eq{\ref{eq:cvxsgd1}}, we require $\eta_t < 1$ and set $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$, 
\begin{align*}
	\sum_{t=0}^{T-1} 2\eta_t (F(\ov{w}_t) - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2)\\ 
    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2)\\ 
    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2G^2 ( 4L(E-1)^2 + 1)\\
    \sum_{t=0}^{T-1} 2\eta_t (F^*_t - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2G^2 ( 4L(E-1)^2 + 1)\\
 F_t^* - F^*  & \leq \frac{\EE \|\ov{w}_{0}-\vw^{*}\|^{2}}{2 \sum_{t=0}^{T-1} \eta_t } + ( 4L(E-1)^2 + 1)G^2 \sum_{t=0}^{T-1} \eta_t^2
\end{align*}
It will converge under three conditions $ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t = \infty$
$ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t^2 < \infty$. For example, we can set $\eta_t = \frac{1}{t+1}$.