### 2020-04-03 source/sec_convexsmoothsgd.tex
partial cvx smth nasgd

% !TEX ROOT=./main.tex

Stochastic gradient of device $k$ at time step $t$, at point $\vw_t^k$: 
	$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$,
	$ \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $,
	$\EE \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}\right)$.
One-step stochastic gradient of all devices:
$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $, $\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}$

\begin{theorem}
Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, we have,
$$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
\label{th:sgdcvxsmth}
\end{theorem}

\begin{proof}

\begin{align}
\EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \vw^*\|^2 &= \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1} + \overline{\vv}_{t+1} - \vw^*\|^2\\
&= \EE_{\cS_{t+1}, \xi_{t}} \left[\|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \|\overline{\vv}_{t+1} - \vw^*\|^2\right] \\
& \hspace{3em}+ 2\EE_{\xi_{t}} \left<\EE_{\cS_{t+1}} \overline{\vw}_{t+1} - \overline{\vv}_{t+1},   \overline{\vv}_{t+1} - \vw^*\right> \\
& = \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \\
& \leq  \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \label{eq:sgdcvxsmth1}
\end{align}
where in the second line we use $\EE_{\cS_{t+1}} \overline{\vw}_{t+1}  = \overline{\vv}_{t+1}$. 
The first term can be bounded by consider different sampling scheme. 
According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand the second term in \eq{\ref{eq:sgdcvxsmth1}} as follows:
$$\begin{aligned}\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*} \right>+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$

Take the expectation condition on $\vw_t$ over $\xi_t$, i.e., random samples at all devices:
\begin{align}
\EE\left[\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\ov{w}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*} \right> +\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
\label{eq:expandsgd}
\end{align}

Now we focus on bounding $-2 \eta_{t} \left< \EE \vg_{t}, \ov{w}_{t}-\vw^{*}\right>$ in \eq{\ref{eq:expandsgd}}:

\begin{align*}
	& -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*}\right> \\
 =  & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> -2 \eta_{t} \left<\EE \vg_{t}, \vw^{k}_t - \vw^{*}\right>\\
 \leq & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\\
 \leq &\sum_{k=1}^N p_k \left[2 \eta_{t} (F_k(\vw^{k}_t) - F_k(\ov{w}_t) + \frac{L}{2} \|\ov{w}_t - \vw^{k}_t\|^2 ) + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\right]\\
 = & \sum_{k=1}^N p_k \eta_t L \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} \sum_{k=1}^N p_k (F_k(w^*) - F_k(\ov{w}_t))\\
 = &  \eta_t L \sum_{k=1}^N p_k \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))
\end{align*}
Plug in this upper bound into \eq{\ref{eq:expandsgd}}, \eq{\ref{eq:sgdcvxsmth1}} and take totoal expectation over all samples at all iterations, we have
\begin{align}
\EE\left\|\ov{w}_{t+1}-\vw^{*}\right\|^2 &\leq \EE\|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \eta_t L \sum_{k=1}^N p_k \EE \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - \EE F(\ov{w}_t)) + \eta_t^2 C\\
&\leq\EE \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - \EE F(\ov{w}_t)) + \eta_t^2 C\\
&\leq \EE\|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} G^2 +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - \EE F(\ov{w}_t))+\eta_t^2 C \label{eq:cvxsgd1}
\end{align}

Sum two sides of \eq{\ref{eq:cvxsgd1}}, we require $\eta_t < 1$ and set $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$, 
\begin{align*}
	\sum_{t=0}^{T-1} 2\eta_t (\EE[F(\ov{w}_t)] - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2+ \eta_t^2 C)\\ 
    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2+ \eta_t^2 C)\\ 
    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2\left[ G^2 ( 4L(E-1)^2 + 1) + C\right]\\
    \sum_{t=0}^{T-1} 2\eta_t (\EE[F^*_t] - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2\left[ G^2 ( 4L(E-1)^2 + 1) + C\right]\\
\EE[ F_t^*] - F^*  & \leq \frac{\|\ov{w}_{0}-\vw^{*}\|^{2}}{2 \sum_{t=0}^{T-1} \eta_t } + \frac{\left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2}{2 \sum_{t=0}^{T-1} \eta_t }
\end{align*}
It will converge under the following conditions: $ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t = \infty$, 
$ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t^2 < \infty$. 
\begin{align*}
	\EE[ F_t^*] - F^*  & \leq \frac{\|\ov{w}_{0}-\vw^{*}\|^{2}}{2 \sum_{t=0}^{T-1} \eta_t } + \frac{\left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2}{2 \sum_{t=0}^{T-1} \eta_t } \\
	\EE[ F_t^*] - F^*  & \leq \frac{\Delta_0 + \left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2 }{2 \sum_{t=0}^{T-1} \eta_t }\\
	& \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}
\end{align*}
where in the last line we set the learning rate satisfying $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, 
and denote $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$ .
\end{proof}