### 2020-05-29 source/sec_strongly_convex_smooth.tex
updated strongly convex smooth section

% !TEX ROOT=./main.tex

\section{Strongly Convex Smooth Objectives}

In this section, we show that FedAvg with local SGD has $O(1/NT)$ convergence
rate for $\mu$-strongly convex and $L$-smooth objectives. The result
improves on the $O(1/T)$ result of {[}ICLR{]} with a linear speedup
in the number of devices $N$. Moreover, it makes a distinction on
the choice of $E$ with partial and full participation. With full
participation, $E$ can be chosen as large as $O(\sqrt{\frac{T}{N}})$
without degrading the linear speedup in the number of workers. On
the other hand, with partial participation, $E$ must be $O(1)$. 
\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
	$k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$ where
	$E$ is the communication delay, and learning rates 
	\begin{align*}
	\alpha_{t} & =\frac{c}{\mu(\gamma+t)}
	\end{align*}
	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
	is such that $\alpha_{t}\leq\frac{1}{8L}$ for all $t\geq0$.

	Then with full device participation, if $\nu_{max}=\max_{k}\frac{1}{N}p_{k}$,
	we have 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\cdot(\frac{\kappa\nu_{max}^{2}\sigma^{2}}{NT}+\frac{5\kappa E^{2}LG^{2}}{T^{2}})
	\end{align*}
	and with partial participation with $K$ sampled devices at each
	communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C(\frac{\kappa\nu_{max}^{2}\sigma^{2}}{NT}+\frac{4\kappa E^{2}G^{2}}{KT}+\frac{5\kappa E^{2}LG^{2}}{T^{2}})
	\end{align*}
\end{theorem}
%
\begin{remark}
	We remark on the linear speedup and the choice of $E$. In full participation,
	the leading term is $\frac{1}{NT}\nu_{max}^{2}\sigma^{2}$ whereas
	$\frac{5\kappa E^{2}LG^{2}}{T^{2}}$ is a lower order term if $E$
	is chosen constant. This implies that $E$ can be chosen to be as
	large as $E=O(\sqrt{\frac{T}{N}})$ which results in $\frac{5\kappa E^{2}LG^{2}}{T^{2}}=O(\frac{1}{NT})$,
	which will be on the same order as the leading term. This is in contrast
	with the convergence rate in {[}ICLR{]} where $E$ cannot scale with
	$T$ and must be $O(1)$, even with full participation. Thus our bound
	yields a more efficient communication complexity than previous analyses.
	We note however that this is only possible when there is full participation,
	and when $\nu_{max}$ is close to 1. When there is partial participation,
	however, $E$ has to stay $O(1)$, as the term leading term in the
	bound for partial participation contains $\frac{4\kappa E^{2}G^{2}}{KT}$.
	Thus our result hightlights the difference between the communication
	efficiency of full and partial participation when an $O(1/NT)$ bound
	on the optimality gap is desired.

	\textbf{}%
	\begin{comment}
	This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
	without degrading the performance. This should be checked in experiments,
	whether with partial participation if the communication round is set
	to scale with $T$, the convergence deteriorates. This is in constrast
	with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
	is allowed. \textbf{If we can confirm that full participation allows
	linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
	whereas partial participation only allows $E=O(1)$, then this would
	an interesting phenomenon that is not reported by previous studies!}
	\end{comment}
\end{remark}

\subsection{Accelerated SGD}
In this section, we show that FedAvg with Nesterov accelerated SGD updates enjoys similar convergence rates with linear speedup.

The FedAv algorithm with Nesterov Accelerated
SGD follows the updates
\begin{align*}
y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align*}
where 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align*}
is the stochastic gradient.

\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
	$k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$ where
	$E$ is the communication interval, and learning rates 
	\begin{align*}
	\alpha_{t} & =\frac{c}{\mu(\gamma+t)}\\
	\beta_{t} & \leq\alpha_{t}
	\end{align*}
	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
	is small enough such that the following hold: 
	\begin{align*}
	\alpha_{t}^{2}+\beta_{t-1}^{2} & \leq\frac{1}{2}\\
	\alpha_{t} & \leq\frac{1}{4L}\\
	4\alpha_{t-1}^{2} & \leq\alpha_{t}
	\end{align*}
	for all $t\geq0$. Suppose also that $G$ is a constant satisfying
	$\mathbb{E}\|w_{0}-\alpha_{0}g_{0,k}\|^{2}=\mathbb{E}\|w_{0}-\alpha_{0}\nabla F_{k}(w_{0},\xi_{0}^{k})\|^{2}\leq G^{2}$
	for all $k$, and $\mathbb{E}\|g_{t,k}\|^{2}\leq G^{2}$ for all $t,k$.

	Then with full device participation, if $\nu_{max}=\max_{k}\frac{1}{N}p_{k}$,
	we have 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(32E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2})
	\end{align*}
	so that as long as $E=O(\sqrt{\frac{N}{T}})$, 
	\begin{align*}
	\mathbb{E}F(w_{T})-F^{\ast} & \leq C(32LG^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{N(\gamma+T)}
	\end{align*}
	and with partial participation with $K$ sampled devices at each
	communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(32E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{16}{K}E^{2}G^{2})
	\end{align*}
	so that if $E=O(1)$, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast} & \leq C(16E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+T)}+\frac{32E^{2}LG^{2}}{T^{2}}
	\end{align*}
\end{theorem}