### 2020-05-30 source/sec_relatedwork.tex
related work v0

% !TEX ROOT=./main.tex

\section{Related Works}

% related work on convergence rate
Federated learning (FL) was originally proposed
in~\cite{mcmahan2016communication} for learning a single, global statistal
model from the isolated data stored in a massive number of devices.  The
empirical success of FL~\cite{chen2018federated,47586} has attracted much
attentions on the rigorous theoretical understanding of the leading algorithm 
in the field: Federated Averaging (FedAvg) and its accelerated variants~\cite{liu2019accelerating,haddadpour2019convergence,khaled2019first,li2019convergence,huo2020faster,yu2019linear,yu2019parallel,koloskova2020unified}.

% Non-convex linear speedup
In particular, we focus on the convergence analysis of FedAvg, which
has been discussed in various settings. 
In \cite{yu2019parallel,wang2019adaptive}, the authors provides convergence
rate of FedAvg for non-convex problem, given the data are i.i.d. distributed.