### 2020-06-02 source/sec_exp.tex
exp cut

% !TEX ROOT=./main.tex

\section{Numerical Experiments}
\label{sec:exp}

In this section, we empirically examine the linear speedup convergence of FedAvg and Nesterov accelerated FedAvg on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as proved in previous theorems.

\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, we
conduct experiments on both synthetic datasets and real-world dataset
w8a~\cite{platt1998fast} $(d=300, n=49749)$. We consider the distributed
objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
$k$-th local device includes three cases: 1) \textbf{Strongly convex
objective}: the regularized binary logistic regression problem, $F_k(\vw) =
\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \frac{\lambda}{2}
\|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
1e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
problem without regularization. $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k}
\log( 1+ \exp(-y_i^k \vw^T\vx_i^k)$. 3) \textbf{Overparameterized setting}:
the linear regression problem without adding noise to the label, $F_k(\vw) =
\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  In the first
two cases, we use the w8a dataset to conduct empirical evaluation. 
In the linear regression problem, we use features from w8a dataset and 
generate a ground truth $w$ to label each instance, which satisfies the
overparameterized setting. For the detailed experimental setup and
hyperparameter setting, please refer to the section~\ref{sec:expsupp} in the
appendix.

% ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
\textbf{Linear speedup.} To verify the linear speedup results of FedAvg in Theorem~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}, we evaluate the number of iterations needed to reach
$\epsilon$-accuracy in three objectives. We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
the best seed results for each configuration $(E, K)$. In each configure, we
extensive search the stepsize from $\min(\eta_0, \frac{nc}{1 + t})$, where
$\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup} (1st row),
the number of iterations to reach the target accuracy $\epsilon$ decreasing 
as the number of workers increase.

\textbf{Partial participations.} We verify the linear speedup in the
partial participation settings, where we set $50\%$ of devices
are active. As the results are shown in Figure~\ref{fig:speedup} (2nd row), the 
FedAve enjoys linear speedup in various settings even with partial
device participation.

\textbf{Nesterov accelerated FedAvg.} In the third row of Figure~\ref{fig:speedup}, 
we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg. The empirical observations align with Theorem~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx} that the accelerated version of FedAvg can also achieve the linear speedup
w.r.t the number of workers.

\textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
affect convergence. As shown in Figure~\ref{fig:e} in appendix section~\ref{sec:expsupp}, the number of iterations increases as $E$ increase, which slow down the convergence in
terms of gradient computation. However, it can save communication costs as
the number of rounds decreased when the $E$ increases. This showcase that
we need a proper choice of $E$ to trade-off the communication cost and
convergence speed.

\begin{figure}[t!]
\centering
	\begin{tabular}{ccc}
	\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
\includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-partialstronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
\includegraphics[width=0.33\textwidth]{fig/paper-partialcvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
\includegraphics[width=0.33\textwidth]{fig/paper-partiallinregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
& 
\includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
	\end{tabular}
\caption{\textbf{Linear speedup} The first row reports the linear speedup convergence of FedAvg w.r.t the number of workers. The second row reports the linear speedup convergence of FedAvg w.r.t the number of active workers in the partial parcipation setting. The third row reports the linear speedup convergence of Nesterov accelerated FedAvg in full participation setting. }
\label{fig:speedup}
\end{figure}

% \begin{figure}
% \centering
% 	\begin{tabular}{ccc}
% 	\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-partialstronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
% \includegraphics[width=0.33\textwidth]{fig/paper-partialcvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
% \includegraphics[width=0.33\textwidth]{fig/paper-partiallinregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
% (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
% 	\end{tabular}
% \caption{The linear speedup convergence of FedAvg w.r.t the number of active workers. }
% \vspace{-2em}
% \label{fig:partial}
% \end{figure}

% \begin{figure}
% \centering
% \begin{tabular}{ccc}
% \hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
% \includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
% & 
% \includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
% (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
% 	\end{tabular}
% \caption{The linear speedup convergence of Nesterov accelerated FedAvg w.r.t the number of workers. }
% \label{fig:nesterov}
% \end{figure}

% \begin{figure}
% \centering
% 	\begin{tabular}{ccc}
% 	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
% 	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
% 	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
% 	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
% 	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsRounds-min-w8a-epsilon0134-reg0.pdf} & 
% 	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsRounds-min-linearregressionw8a-epsilon002-reg0.pdf} \\
% (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
% 	\end{tabular}
% \caption{The convergence of FedAvg w.r.t the number of local steps $E$. }
% \label{fig:e}
% \end{figure}