### 2020-06-03 source/sec_overparameterized.tex
Sec 5
1.  Shrink the summary in first paragraph.
2.  Move discussion of related work on accelerated FedAvg from intro to 5.3.
3. remove some computation in L244.

% !TEX ROOT=./main.tex

\section{Geometric Convergence of FedAvg in the Overparameterized Setting}
\label{sec:overparameterized}

% In this section, we turn to the special setting of overparameterized
% problems, where a non-negative loss function admits a zero-loss global
% minimizer. Such problems are prevalent especially at the age of deep learning,
% where the number of parameters far exceeds the number of data points,
% and perfect fitting is possible. There have been a line of recent
% works showing that SGD and accelerated methods achieve geometric
% convergence in similar settings thanks to the property of \textit{automatic
% variance reduction}~\cite{ma2017power,moulines2011non,needell2014stochastic,schmidt2013fast,strohmer2009randomized}.
% The natural question is whether such a result also holds in the federated
% learning setting. 
% We show that this is indeed the case and establish
% the geometric (linear) convergence of FedAvg with local SGD updates
% with constant step size for general strongly convex and smooth overparameterized
% problems. In addition, we show that the convergence rate speeds up
% linearly in the number of workers $N$ when $N$ is below some problem-dependent
% threshold, while it decreases with $E$ through $1/E$.

Overparameterization is a prevalent machine learning setting where the
statistical model has much more parameters than the number of training samples
and the existence of parameter choices with zero training loss is
assured~\cite{allen2018convergence,zhang2016understanding}.  Due to the
property of \textit{automatic variance reduction} in overparameterization,  a
line of recent works proved that SGD and accelerated methods achieve geometric
convergence~\cite{ma2017power,moulines2011non,needell2014stochastic,schmidt2013fast,strohmer2009randomized}.
A natural question is whether such a result still holds in the federated
learning setting. 
In this section, we provide the first geometric convergence rate
of FedAvg for the overparameterized strongly convex and smooth problems,
and we show that it preserves the linear speedup at the same time.  
We then sharpen this result in the special case of
quadratic problems.
Inspiring by the recent advances in SGD~\cite{liu2019accelerating,jain2017accelerating}, 
we further propose a novel momentum-based FedAvg algorithm, which enjoys an
improved convergence rate over FedAvg. 
\begin{comment}
we deal with the question of whether FedAvg with momentum-based
local updates can outperform FedAvg with SGD updates. In contrast
to the gradient descent setting, Nesterov and Heavy Ball updates are
known to fail to accelerate over SGD, both in the overparameterized
setting and standard convex setting \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
Thus in general one cannot hope to obtain acceleration results for
the FedAvg algorithm with Nesterov and Heavy Ball updates. On the
hopeful side, \cite{jain2017accelerating,liu2018accelerating} introduced
similar algorithms that make modifications of the Nesterov updates
to correct for its ``over-descent''. For quadratic objectives in
the overparamterized setting, \cite{liu2018accelerating} show their
algorithm achieves acceleration over the geometric convergence of
SGD that recovers the well-known $\mathcal{O}(\exp(-t/\sqrt{\kappa})$
acceleration of Nesterov over GD. In the last part, we introduce a
new accelerated FedAvg algorithm by adapting the MaSS algorithm of
\cite{liu2018accelerating} to the federated learning setting. We
show that it achieves geometric convergence for overparameterized
quadratic problems with rate $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
where $\tilde{\kappa}$ is a ``statistical condition number''\cite{liu2018accelerating,jain2017accelerating}
that satisfies $\tilde{\kappa}\leq\kappa_{1}$. Thus our new FedAvg
algorithm achieves a speedup of factor $\sqrt{\kappa_{1}/\tilde{\kappa}}$
over FedAvg with local SGD updates. 
\end{comment}

\subsection{Geometric Convergence of FedAvg in the Overparameterized Setting}
Recall the federated learning problem $\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
with $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(\mathbf{w};\mathbf{x}_{k}^{j})$.
In this section, we consider the standard Empirical Risk Minimization (ERM) setting where $\ell$
is non-negative, $l$-smooth, and convex, and as before, we assume
that each $F_{k}(\mathbf{w})$ is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
setup includes many important problems in practice. In the overparameterized
setting, there exists $\mathbf{w}^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
such that $\ell(\mathbf{w}^{\ast};\mathbf{x}_{k}^{j})=0$ for all
$\mathbf{x}_{k}^{j}$. We first show that FedAvg with local SGD updates
and communication every $E$ iterations achieves geometric convergence
with linear speedup in the number of workers. 
\begin{theorem}
	\label{thm:overparameterized_general}In the overparameterized setting,
	FedAvg with local SGD updates and communication every $E$ iterations
	with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
	has geometric convergence,
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp(-\frac{\mu}{E}\frac{NT}{l\nu_{\max}+L(N-\nu_{\min})})\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
	\end{align*}
\end{theorem}
%
\textbf{Linear speedup and Communication Complexity} When $l>L$,
the speedup factor is on the order of $\mathcal{O}(N/E)$ %\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=
for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
every $E$ iterations provides a geometric convergence speedup factor
of $\mathcal{O}(N/E)$, for $N\leq\frac{l}{L}+1$. When $N$ is above
this threshold, however, the speedup is almost constant in the number
of workers. This matches the findings in \cite{ma2017power}. Our
result also illustrates that $E$ can be taken $\mathcal{O}(T^{\beta})$
for any $\beta<1$, achieving better communication efficiency than
the standard setting.

\subsection{Overparameterized Quadratic Problems}
We now turn to quadratic problems
and demonstrate how the general bound in Theorem~\ref{thm:overparameterized_general}
can be sharpened to explicitly depend on the condition number of the
system, and how this rate can be further improved with a new FedAvg
algorithm with momentum updates.

The local device objectives are now given by the sum of squares {\small$F_{k}(\mathbf{w})=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{w}^{T}\mathbf{x}_{k}^{j}-z_{k}^{j})^{2}$},
and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})\equiv0$.
Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$,
the averaged Hessian matrix as $H:=\sum_{k=1}^{N}p_{k}H^{k}$, and the stochastic Hessian matrix as $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$. 
\begin{comment}
In general $H$ has zero eigevalues. However, because the null space
of $H$ and range of $H$ are orthogonal, in our subsequence analysis
it suffices to project $\overline{\mathbf{w}}_{t}-\mathbf{w}^{\ast}$
onto the range of $H$, thus we may restrict to the non-zero eigenvalue
of $H$. We can use $\mathbf{w}^{\ast T}\mathbf{x}_{k}^{j}-z_{k}^{j}\equiv0$
to rewrite the local objectives as $F_{k}(\mathbf{w})=\frac{1}{2}\langle\mathbf{w}-\mathbf{w}^{\ast},H^{k}(\mathbf{w}-\mathbf{w}^{\ast})\rangle\equiv\frac{1}{2}\|\mathbf{w}-\mathbf{w}^{\ast}\|_{H^{k}}^{2}$
so that $F(\mathbf{w})=\frac{1}{2}\|\mathbf{w}-\mathbf{w}^{\ast}\|_{H}^{2}$.
\end{comment}
\begin{comment}
\begin{align*}
F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j}-(w^{\ast T}x_{k,j}-z_{k,j}))^{2}=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}((w-w^{\ast})^{T}x_{k,j})^{2}\\
& =\frac{1}{2}\langle w-w^{\ast},H^{k}(w-w^{\ast})\rangle=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}
\end{align*}
\end{comment}
\begin{comment}
Let $\xi_{t}^{k}$ be the stochastic sample on the $k$th device at
time $t$, and define $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$
as the stochastic Hessian matrix. %	
\end{comment}
\begin{comment}
Note that $\mathbb{E}\tilde{H}_{t}^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}=H^{k}$
and $\mathbf{g}_{t,k}=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})=\tilde{H}_{t}^{k}(\mathbf{w}_{t}^{k}-\mathbf{w}^{\ast})$
while $\mathbf{g}_{t}=\sum_{k=1}^{N}p_{k}\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})=\sum_{k=1}^{N}p_{k}\tilde{H}_{t}^{k}(\mathbf{w}_{t}^{k}-\mathbf{w}^{\ast})$.
\end{comment}
Define $l$ to be the smallest positive number such that $\mathbb{E}\|\xi_{t}^{k}\|^{2}$$\mathbf{\xi}_{t}^{k}$($\mathbf{\xi}_{t}^{k})^{T}\preceq lH^{k}$
for all $k$. Note that $l\leq\max_{k,j}\|\mathbf{x}_{k}^{j}\|^{2}$.
Let $L$ and $\mu$ be lower and upper bounds of non-zero eigenvalues
of $H^{k}$. Define $\kappa_{1}:=l/\mu$ and $\kappa:=L/\mu$. Following
\cite{liu2018accelerating,jain2017accelerating}, we define the statistical
condition number $\tilde{\kappa}$ as the smallest positive real number
such that $\mathbb{E}\left[\langle\xi_{t}^{k}(H^{k})^{-1},\xi_{t}^{k}\rangle\xi_{t}^{k}(\xi_{t}^{k})^{T}\right] \  \preceq\tilde{\kappa}H^{k}$, for all $k$. 
The condition numbers $\kappa_{1}$ and $\tilde{\kappa}$
are important in the characterization of convergence rates for FedAvg
algorithms. Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.

\subsection{Acceleration of FedAvg with Momentum-based Updates}

We show that the bound in \ref{thm:overparameterized_general} can
be improved for quadratic problems with explicit dependence on condition
number, which is then used to illustrate the acceleration of our proposed
FedAvg with MaSS updates over SGD updates.
\begin{theorem}
	\label{thm:overparameterized_quadratic}For the overparamterized quadratic
	problem, FedAvg with local SGD updates and communication every $E$
	iterations with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})})$
	has geometric convergence:
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\mathcal{O}\left(L(1-\frac{N}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
	\end{align*}
\end{theorem}

When $N=\mathcal{O}(\kappa_{1})$, the convergence rate is $\mathcal{O}((1-\frac{N}{E\kappa_{1}})^{T})=\mathcal{O}(\exp(-\frac{NT}{E\kappa_{1}}))$,
which exhibits linear speedup in the number of workers, as well as
a $1/\kappa_{1}$ dependence on the condition number $\kappa_{1}$.
Inspired by \cite{liu2018accelerating}, we propose the \textbf{MaSS
	accelerated FedAvg algorithm} (FedMaSS):
\begin{align*}
\mathbf{w}_{t+1}^{k} & =\begin{cases}
\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k} & \text{if }t+1\notin\mathcal{I}_{E},\\
\sum_{k=1}^{N}p_{k}\left[\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E},
\end{cases}\\
\mathbf{u}_{t+1}^{k} & =\mathbf{w}_{t+1}^{k}+\gamma^{k}(\mathbf{w}_{t+1}^{k}-\mathbf{w}_{t}^{k})+\eta_{2}^{k}\mathbf{g}_{t,k}. 
\end{align*}
When $\eta_{2}^{k}\equiv0$, this algorithm reduces to the Nesterov
accelerated FedAvg algorithm. In the next theorem, we demonstrate
that FedMaSS improves the convergence to $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$.
To our knowledge, this is the first accelerated result of
FedAvg with momentum updates over SGD updates.
\begin{theorem}
	\label{thm:overparameterized_MaSS}For the overparamterized quadratic
	problem, FedMaSS with communication every $E$ iterations and constant
	step sizes $\overline{\eta}_1=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}),\overline{\eta}_{2}=\frac{\overline{\eta}_{1}(1-\frac{1}{\tilde{\kappa}})}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}},\overline{\gamma}=\frac{1-\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}$
	has geometric convergence:
	\begin{align*}
	\mathbb{E}F(\overline{\vw}_{T}) & \leq\mathcal{O}\left((1-\frac{1}{E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{T}\|w_{0}-w^{\ast}\|^{2}\right).
	\end{align*}
\end{theorem}
\textbf{Speedup of FedMaSS over FedAvg} 
To better understand the significance of the above result, we briefly discuss the related works in SGD.
Nesterov and Heavy Ball updates are known to fail to accelerate over SGD in both the overparameterized
and convex settings~\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
Thus in general one cannot hope to obtain acceleration results for
the FedAvg algorithm with Nesterov and Heavy Ball updates. 
Luckily, recent works in SGD~\cite{jain2017accelerating,liu2018accelerating} introduced
an additional compensation term to the Nesterov updates to address the non-acceleration issue. Surprisingly, 
we show the same approach can effectively improve the rate of FedAvg.
Comparing the convergence rate of FedMass (Theorem~\ref{thm:overparameterized_MaSS}) and FedAvg (Theorem~\ref{thm:overparameterized_quadratic}), when $N=\mathcal{O}(\sqrt{\kappa_{1}\tilde{\kappa}})$,
the convergence rate is $\mathcal{O}((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{T})=\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$
as opposed to $\mathcal{O}(\exp(-\frac{NT}{E\kappa_{1}}))$. Since
$\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of
$\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for FedMaSS. On the other
hand, the same linear speedup in the number of workers holds for $N$
in a smaller range of values.