### 2020-06-03 source/app_sumarization.tex
add assumption discussion

% !TEX ROOT=./main.tex

\section{Comparison of convergence rates with related works.}
\label{sec:app:comparison}
In this section, we summarize the convergence results in this work from various perspectives
and compare our results with related works. In Table~\ref{tb:convergenceratev1} and Table~\ref{tb:convergenceratev2}, we summarize 
the convergence results in terms of participations and accelerations. In Table~\ref{tb:convergenceratev3}, we compare our 
results with the best known results in the literatures.

In~\cite{haddadpour2019convergence}, they also provide $\cO(1/NT)$ convergence rate of non-convex problems under Polyak-Łojasiewicz (PL) condition, which
means their results can directly apply to the strongly convex problems. However, their assumption is based on bounded gradient diversity, defined as follows: 
\begin{align*}
	\Lambda(\vw) = \frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}} \leq B
\end{align*} 
This is a more restrictive comparing to assuming bounded gradient under the case of target accuracy $\epsilon \rightarrow 0$ and PL condition.
To see this, consider the gradient diversity at the global optimal $\vw^*$,i.e., $\Lambda(\vw^*) = \frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$. For $\Lambda(\vw^*)$ to be bounded, it requires $\|\nabla F_{k}(\mathbf{w}^*)\|_{2}^{2} = 0$, $\forall \ k$. This indicates 
$\vw^*$ is also the minimizer of each local objective, which contradicts to the practical setting of heterogeneous data. Therefore, their bound 
is not effective for arbitrary small $\epsilon$-accuracy under general heterogeneous data while our convergence results still hold in this case.
% If we only consider the domain $\vw \in \{\vw | \|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2} > c\}$, where $c$ is a positive constant. 
% The bounded gradient can imply gradient diversity (see Corollary 10 in ~\cite{li2018federated}). In this case, the gradient diversity is a relaxed

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|}\hline 
	Algorithm \textbackslash\  Objective function        & Strongly Convex        & Convex \\ \hline \hline
	FedAvg                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
	Nesterov accelerated FedAvg    & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    & $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
\end{tabular}
\caption{A high level summary of the convergence results for full participation in this paper. This table only highlights the
dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.}
\label{tb:convergenceratev1}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|}\hline 
	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
\end{tabular}
\caption{A high level summary of the convergence results in this paper. This table only highlights the
dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.
These convergence rates applied to both FedAvg and Nesterov accelerated FedAvg.}
\label{tb:convergenceratev2}
\end{table}

\begin{table}[h!]
\centering
{\tiny
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline Reference                 & Convergence rate    & E                           			& NonIID & Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    			& \cmark & Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^{\dagger}$     	& \cmark$^{\ddagger\ddagger}$ & Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$& \cmark			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
\cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			 & \cmark& Full        & Bounded gradient             & Convex  \\\hline
This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   & \cmark&  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline
This work & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$ & \cmark &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
\end{tabular}
}
\caption{A high level summary of the convergence results in this paper and their comparison to prior state-of-the-art FL algorithms, considering heterogeneous data. This table only highlights the
dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices. 
$\kappa$ is the condition number of the system and $\beta \in (0,1)$. 
We note that all our results for strongly convex and convex smooth problems are unified results for both FedAvg and accelerated FedAvg. The convergence results for prior arts only apply for FedAvg.}
{\raggedright 
         $^{\dagger}$ This $E$ is obtained under i.i.d. setting. \\
         $^{\dagger\dagger}$ This $E$ is obtained under full participation setting. \\ 
         $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.\\
         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition, which also applies for strongly convex problems. Therefore, we compare it with our strongly convex results here. \\
         $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).
           \par}
\label{tb:convergenceratev3}
\end{table}