### 2020-06-03 source/sec_impact.tex
add covid impact

% !TEX ROOT=./main.tex

% Broader impact: Authors are asked to include a section in their submissions discussing the broader impact of their work, including possible societal consequences — -both positive and negative.
\section{Broader impact}
With the widespread use of mobile devices, the powerful sensors
collected enormous amount of data empowered with machine learning
solutions can transform people's life in many aspects. The proper
data integration can produce positive social
impact in terms of better personalization~\cite{fallah2020personalized}, more efficient information solicitation~\cite{chen2018federated}, and enhanced service~\cite{47586}.

In the meantime, the increasing concerns of security and
privacy leaking from both large companies and the entire society
present significant challenges of proper utilization of local
data. This trade-off between efficiency and privacy/security
lies in the center of many industrial applications, including 
but not limited to recommendation systems~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and mobile keyboard prediction~\cite{47586}.

Federated learning balancing efficiency and privacy is not only an effective
tool to provide enhanced entertainment, but also a vital approach that can
save millions of lives in the short future.  Under the unprecedented outbreak
of Covid'19, coronavirus patients have filled the emergency rooms
thoroughly. It is imperative to have reliable, robust machine learning
solutions that could help doctors treat patients. However,  HIPAA regulations
impose strict restrictions on sharing patients' clinical data: privacy is a
top priority when developing machine learning techniques at those medical institutions. 
Therefore, propelling the frontiers of federated learning for more efficient 
algorithms under the privacy-preserving constraint has never been more
important.

In this work, we establish solid theoretical foundations for federated
learning that achieves efficiency and privacy at the same time. 
On one hand, the data is isolated in each local devices, thus prevent
the potential leakage of private information. On the other hand, the
convergence rates can be improved substantially if the number of
active local devices increases. In conclusion, our work 
shows a faithful pathway to simultaneously secure people's privacy and improve
users' experience.