### 2020-06-09 source/app_nasgd_scvx_smooth.tex
updated nesterov proofs

\begin{thm}
	Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$. Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
	\end{align*}
\end{thm}
%
%
\textbf{}%
\begin{comment}
This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
without degrading the performance. This should be checked in experiments,
whether with partial participation if the communication round is set
to scale with $T$, the convergence deteriorates. This is in constrast
with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
is allowed. \textbf{If we can confirm that full participation allows
linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
whereas partial participation only allows $E=O(1)$, then this would
an interesting phenomenon that is not reported by previous studies!}

The convergence result implies that with fixed $E$, as long as \$N\$
satisfies \$E=O(\textbackslash sqrt\{T/N\})\$, there is linear speedup.
When \$N\$ exceeds an upper bound, however, linear speedup may fail
to happen, and \$T/N\$ may remain constant, i.e. the number of iterations
required for convergence may increase with \$N\$. 
\end{comment}

\begin{proof}
Define the virtual sequences $\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$,
$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{v}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
and $\overline{w}_{t+1}=\overline{v}_{t+1}$ for all $t$. The proof
again uses the $L$-smoothness of $F$ to bound 
\begin{align*}
\mathbb{E}(F(\overline{v}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{v}_{t})-F(w^{\ast}))\\
& \leq\frac{L}{2}\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}
\end{align*}
Our main step is to prove the bound
\begin{align*}
\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+20E^{2}L\alpha_{t}^{3}G^{2}
\end{align*}
 for appropriate step sizes $\alpha_{t},\beta_{t}$.

We have 
\begin{align*}
\|\overline{v}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
& =A_{1}+A_{2}+A_{3}
\end{align*}
where 
\begin{align*}
A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
\end{align*}
$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
while for $A_{3}$ we have
\begin{align*}
\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
\end{align*}
again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

Next we bound $A_{1}$: 
\begin{align*}
\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
\end{align*}
Same as the SGD case, 
\begin{align*}
-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
\end{align*}
so that 
\begin{align*}
\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
\end{align*}

Different from the SGD case, we have 
\begin{align*}
\|\overline{w}_{t}-w^{\ast}\|^{2} & =\|\overline{v}_{t}+\beta_{t-1}(\overline{v}_{t}-\overline{v}_{t-1})-w^{\ast}\|^{2}\\
& =\|(1+\beta_{t-1})(\overline{v}_{t}-w^{\ast})-\beta_{t-1}(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
& =(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\overline{v}_{t}-w^{\ast},\overline{v}_{t-1}-w^{\ast}\rangle+\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
& \leq(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|+\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}
\end{align*}
which gives 
\begin{align*}
\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\overline{v}_{t-1}-w^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
\end{align*}
and we will using this recursive relation to obtain the desired bound.

First we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
Since communication is done every $E$ steps, for any $t\geq0$, we
can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
all $k$. Moreover, using $\alpha_{t}$ is non-increasing, $\alpha_{t_{0}}\leq2\alpha{}_{t}$,
and $\beta_{t}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$, we have
\begin{align*}
\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(v_{i+1}^{k}-v_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(v_{i+1}^{k}-v_{i}^{k})\|^{2}\\
& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|g_{i,k}\|^{2}+\|(v_{i+1}^{k}-v_{i}^{k})\|^{2})\\
& \leq4\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}G^{2}\\
& \leq4(E-1)^{2}\alpha_{t_{0}}^{2}G^{2}\leq16(E-1)^{2}\alpha_{t}^{2}G^{2}
\end{align*}
where we have used $\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}\leq G^{2}$.
To see this identity for appropriate $\alpha_{t},\beta_{t}$, note
the recursion 
\begin{align*}
v_{t+1}^{k}-v_{t}^{k} & =w_{t}^{k}-w_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
w_{t+1}^{k}-w_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(v_{t+1}^{k}-v_{t}^{k})
\end{align*}
so that 
\begin{align*}
v_{t+1}^{k}-v_{t}^{k} & =-\alpha_{t-1}g_{t-1,k}+\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
& =\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-\alpha_{t}g_{t,k}
\end{align*}
Since the identity $v_{t+1}^{k}-v_{t}^{k}=\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-\alpha_{t}g_{t,k}$
implies 
\begin{align*}
\mathbb{E}\|v_{t+1}^{k}-v_{t}^{k}\|^{2} & \leq2\beta_{t-1}^{2}\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}+2\alpha_{t}^{2}G^{2}
\end{align*}
as long as $\alpha_{t},\beta_{t-1}$ satisfy $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$,
we can guarantee that $\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}\leq G^{2}$
for all $k$ by induction. This together with Jensen's inequality
also gives $\mathbb{E}\|\overline{v}_{t}-\overline{v}_{t-1}\|^{2}\leq G^{2}$
for all $t$.

Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
we can conclude that, with $\nu_{\max}:=N\cdot\max_{k}p_{k}$,

\begin{align*}
\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|\\
& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|
\end{align*}
where $\sigma^{2}=\sum_{k}p_{k}\sigma_{k}^{2}$, and $\alpha_{t}$
satisfies $L\alpha_{t}\leq\frac{1}{5}$. We show next that $\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$
by induction.

Assume that we have shown 
\begin{align*}
\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2} & \leq b(C\alpha_{t}^{2}+D\alpha_{t})
\end{align*}
for all iterations until $t$, where $C=20E^{2}LG^{2}$, $D=\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
and $b$ is to be chosen later. For step sizes we choose $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$
and $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$
where $\gamma=\max\{32\kappa,E\}$, so that $\beta_{t-1}\leq\alpha_{t}$
and 
\begin{align*}
(1-\mu\alpha_{t})(1+14\beta_{t-1}) & \leq(1-\frac{6}{t+\gamma})(1+\frac{3}{(t+\gamma)(1-\frac{6}{t+\gamma})})\\
& =1-\frac{6}{t+\gamma}+\frac{3}{t+\gamma}=1-\frac{3}{t+\gamma}=1-\frac{\mu\alpha_{t}}{2}
\end{align*}
Recall that we also require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any
$t-t_{0}\leq E-1$, $L\alpha_{t}\leq\frac{1}{5}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$,
which we can also check to hold by definition of $\alpha_{t}$ and
$\beta_{t}$.

Moreover, $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
with the chosen step sizes also implies $\|\overline{v}_{t-1}-w^{\ast}\|\leq2\|\overline{v}_{t}-w^{\ast}\|$.
Therefore the bound for $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}$
can be further simplified with 
\begin{align*}
2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|^{2}
\end{align*}
and 
\begin{align*}
(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t}-w^{\ast})\|^{2}
\end{align*}
so that
\begin{align*}
\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\mathbb{E}\|(\overline{v}_{t}-w^{\ast})\|^{2}\\
& +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\overline{v}_{t}-w^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
& \leq b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
& =(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2}C+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}D
\end{align*}
and so it remains to choose $b$ such that 
\begin{align*}
(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}\\
(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & \leq b\alpha_{t+1}^{2}
\end{align*}
from which we can conclude $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}\leq\alpha_{t+1}^{2}C+\alpha_{t+1}D$.

With $b=\frac{6}{\mu}$, we have
\begin{align*}
(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{6}{\mu(t+\gamma)}\\
& \leq b\frac{6}{\mu(t+\gamma+1)}=b\alpha_{t+1}
\end{align*}
where we have used $\frac{t+\gamma-1}{(t+\gamma)^{2}}\leq\frac{1}{t+\gamma+1}$.

Similarly 
\begin{align*}
(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
& =b(\frac{t+\gamma-2}{t+\gamma})(\frac{6}{\mu(t+\gamma)})^{2}\\
& \leq b\frac{36}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
\end{align*}
where we have used $\frac{t+\gamma-2}{(t+\gamma)^{3}}\leq\frac{1}{(t+\gamma+1)^{2}}$.

Finally, to ensure $\|v_{0}-w^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
we can rescale $b$ by $c\|v_{0}-w^{\ast}\|^{2}$ for some $c.$ It
follows that $\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
for all $t$. Thus

\begin{align*}
\mathbb{E}(F(\overline{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-w^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
& =3c\|w_{0}-w^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
& \leq3c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
\end{align*}

With partial participation, the same argument in the SGD case yields
\begin{align*}
\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
\end{align*}
\end{proof}