### 2020-06-10 source/app_notations.tex
sec A, B, E

% !TEX ROOT=./main.tex

\section{Additional Notations}
\label{sec:app:notations}
In this section, we introduce additional notations that are used throughout
the proof. Following common practice~\cite{stich2018local,li2019convergence}, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
$\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$. For partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\ov{w}_t=\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$. However, we can
set unbiased sampling strategy such that $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$.
$\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. 
\begin{align}
\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t},	\label{eq:vbar}
\end{align}
% $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is one-step stochastic gradient, averaged over all devices. 
\begin{align*}
\vg_{t,k} = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right),   
\end{align*}
% \begin{align}
% \vg_{t,k} \left\{\begin{array}{ll} 
%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
%  \in \partial F_{k}\left(\vw_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth} \lkxcom{remove non-smooth}
%  \end{array}\right.
% \end{align}
Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
\begin{align*}
\EE_{\xi_{t}^{k}} \vg_{t,k}  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right), 
\end{align*}
% \begin{align}
% \EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
%  \in \partial F_{k}\left(\vw_{t}^{k}\right)  & \text{non-smooth}
%  \end{array}\right.
% \end{align}
and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$. 
Since in this work, we also consider the case of partial participation. The sampling
strategy to approximate the system heterogeneity can also affect the convergence. Here we
follow the prior arts~\cite{haddadpour2019convergence} considering two types of sampling
schemes. 
The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by~\cite[Lemma 5]{li2019convergence}:
\begin{align}
\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2.
\end{align}
The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
replacement, in which we have the 
\begin{align}
\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2.
\end{align}
We denote this upper bound as follows for concise presentation. 
\begin{align}
	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C.
	\label{eq:partialsample}
\end{align}