### 2020-08-07 rebuttal/main.tex
rebutal template

\documentclass{article}

\usepackage{neurips_2019_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{color}
\usepackage{lipsum}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}

\input{bmacros}
\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
\begin{document}
The authors would like to thank three reviewers for their valuable feedback. 
% Based on their comments, upon acceptance of this paper, we shall improve in 
% the final version (a) Refine the presentation, fix typo, figures, and properly 
% located assumptions etc. (b) More experiments in terms of 
Our point-to-point response are provided below.

\input{r1}
\input{r2}
\input{r3}

\end{document}