### 2020-08-07 rebuttal/r3.tex

{\color{blue}\textbf{Reviewer 3:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows: \\
\textbf{\textit{1.Comparison with SCAFFOLD.}}: We thank the reviewer point out 
relavent reference, we will cite the paper in the new version.

\textbf{\textit{2. The convergence of Nestrov Accelerated FedAvg.}}

\textbf{\textit{3. The geometrical rates of overparametrized linear regression.}}