### 2020-08-10 rebuttal/acresponse.tex
Update on Overleaf.

\documentclass{article}

\usepackage{neurips_2020_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{color}
\usepackage{lipsum}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{enumitem}
\setlist{leftmargin=2mm}
\input{bmacros}
\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
\begin{document}
Dear Area Chair,

We would like to emphasize that we respectfully disagree with the evaluation from R1 and R3. 
R1 mainly challenge the writing, contributions, and comparison with prior work. Regarding the writing, the other reviewers think the paper is well-written and is easy to follow. Regarding the contributions and comparison with prior work, we explicitly summarized those in the table and did thorough comparison in the appendix.  We also explicitly mention what is the improvement leads to the linear speedup comparing to prior prove in the appendix. 
R1 might think our work is an extension of [21]. This is correct for the strongly convex case. (and we did specify the non-trivial step compared to [21] in the appendix (L483, L492). However,  the general convex and overparameterized setting as well as the Nesterov results use very different techniques and it is not fair to require a simple highlight of non-trivial step compared to prior work [20], as discussed in Appendix B, we use a more general assumption than [20].

R3 mentioned a missing reference (SCAFFOLD) while his/her conclusions of that paper are not correct. The missing reference’s analysis for FedAvg (c.f. Theorem V) didn’t achieve linear speedup in general, as we discuss in detail in our response to R3. On the other hand, our results achieve linear speedup under more flexible sampling strategies for any number of participating devices. Also,  SCAFFOLD didn’t achieve better communication complexity as R3 claimed. We have the same communication complexity and we guarantee the linear speedup.

In light of the above discussions, we kindly request the area chair to re-consider the evaluation from R1 and R3.

\end{document}