### 2020-09-09 source-iclr/sec_intro_old.tex

% !TEX ROOT=./main.tex

\section{Introduction and Related Works}

\begin{table}[h!]
	\centering
	\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         &  Cvx Non-smth & Cvx L-smth & Strongly Cvx Non-smth& Strongly Cvx L-smth \\ \hline
		Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{KT}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{KT}+\frac{E^2}{T^2})$    \\\hline
		SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{KT}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{KT}+\frac{E^2}{T^2})$ ours    \\\hline
	\end{tabular}
	\caption{Summary of related work on federated learning, heterogeneous data.}
\end{table}

+ exponential acceleration of MaSS for linear regression: for $E=1$, 
\begin{align*}
\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N})(\tilde{\kappa}_{N})}})\leq C\cdot\exp(-\frac{Nt}{\sqrt{(\kappa_{1})(\tilde{\kappa})}})
\end{align*}

Most recent result by Koloskova et. al gives same bounds for convex and strongly convex.