### 2020-09-25 source-iclr/app_unifiedframework.tex
Update on Overleaf.

In this section, we provide a unified proof framework for FedAvg and its accelerated variants.