### 2021-07-04 source-nips/MaSS_FedAvg_1.tex
Update on Overleaf.

%% LyX 2.3.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{algorithm,algpseudocode}

\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}

\makeatother

\usepackage{babel}
\providecommand{\theoremname}{Theorem}

\begin{document}

\section{Exponential Convergence of FedAvg with MaSS for Linear Regression
Problem with 0 Loss}

The MaSS algorithm is a variant of the Nesterov acceleration, where
in the update a non-negative multiple of the gradient is added to
the Nesterov parameter update, to correct for the ``over-descent''
of the Nesterov update. In {[}Liu\&Belkin{]}, the authors show that
for stochastic gradient methods in the interpolation setting, Nesterov
update is not able to achieve accleration over vanilla SGD, whereas
the MaSS algorithm is able to achieve acceleration both theoretically
and empirically. We adapt the MaSS algorithm to our federated learning
setting, and prove that for certain class of convex objectives with
0 global objective value, the FedAvg with MaSS update achieves exponential
convergence and acceleration over SGD.

We first show that FedAvg with MaSS has exponential convergence for
linear regression when the global objective has 0 as its global minimum.
Since the global loss is given by 
\begin{align*}
F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
\end{align*}
 and there exists $w^{\ast}$ such that $F(w^{\ast})=0$, in particular
this implies that $F_{k}^{\ast}=F_{k}(w^{\ast})=F^{\ast}=0$ for all
$k$.

\subsection{Notation and Definitions}

Following Liu\&Belkin and Jain et al., we define some condition number
related quantities. Let $H^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}x_{k,j}$
be the Hessian matrix of $F_{k}$. Let $L^{k}$ and $\mu^{k}$ be
the largest and smallest non-zero eigenvalues of the Hessians $H^{k}$.
For a mini-batch $\{\tilde{x}_{j}\}_{j=1}^{m}$ of $m$ samples from
device $k$, let $\tilde{H}_{m}^{k}=\frac{1}{m}\sum_{j=1}^{m}\tilde{x}_{j}\tilde{x}_{j}^{T}$
be the unbiased mini-batch estimate of $H^{k}$. Let $L_{1}^{k}$
be the smallest positive number such that 
\begin{align*}
\mathbb{E}\left[\|\tilde{x}\|^{2}\tilde{x}\tilde{x}^{T}\right] & \preceq L_{1}^{k}H^{k}
\end{align*}
 and define 
\begin{align*}
L_{m}^{k} & =L_{1}^{k}/m+(m-1)L/m
\end{align*}

Define the $m$-stochastic condition number as $\kappa_{m}^{k}:=L_{m}^{k}/\mu$.
Define the statistical condition number $\tilde{\kappa}^{k}$ as the
smallest positive real number such that 
\begin{align*}
\mathbb{E}\left[\langle\tilde{x}(H^{k})^{-1},\tilde{x}\rangle\tilde{x}\tilde{x}^{T}\right] & \preceq\tilde{\kappa}^{k}H^{k}
\end{align*}
 Note that $L_{m}^{k}\geq L$, and $\kappa_{m}^{k}\geq\kappa^{k}$,
while $\tilde{\kappa}^{k}\leq\kappa^{1}$.

\subsection{FedAvg with MaSS}

The FedAvg algorithm with MaSS follows the updates
\begin{align*}
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta_{1}^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta_{1}^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =w_{t+1}^{k}+\gamma^{k}(w_{t+1}^{k}-w_{t}^{k})+\eta_{2}^{k}g_{t,k}
\end{align*}
 where we note that the natural parameter is $w_{t}$, while $u_{t}$
is an auxiliary parameter, which we initialize to be $u_{0}^{k}\equiv u_{0}$,
and 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align*}
 is the stochastic gradient and 
\begin{align*}
g_{t} & =\sum_{k=1}^{N}p_{k}g_{t,k}
\end{align*}
is the averaged stochastic gradient. When $\eta_{2}^{k}\equiv0$,
this reduces to the FedAvg algorithm with Nesterov updates.

We note that the update can equivalently be written as 
\begin{align*}
v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}w_{t}^{k}-\delta^{k}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
 where there is a bijection between the parameters 
\begin{align*}
\frac{1-\alpha^{k}}{1+\alpha^{k}} & =\gamma^{k}\\
\eta^{k} & =\eta_{1}^{k}\\
\frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}} & =\eta_{2}^{k}
\end{align*}
 and we further introduce an auxiliary parameter $v_{t}^{k}$, which
is initialized at $v_{0}^{k}\equiv v_{0}$. We also note that when
$\delta_{t}=\frac{\eta_{t}}{\alpha_{t}}$, the update reduces to the
Nesterov accelerated SGD. This version of the FedAvg with MaSS algorithm
is used for analyzing the exponential convergence.

As before, define the virtual sequences $\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$,
$\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$, $\overline{u}_{t}=\sum_{k=1}^{N}p_{k}u_{t}^{k}$,
and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$. We have
$\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{w}_{t+1}=\overline{u}_{t}-\eta_{t}g_{t}$,
$\overline{v}_{t+1}=(1-\alpha^{k})\overline{v}_{t}+\alpha^{k}\overline{w}_{t}-\delta^{k}g_{t}$,
and $\overline{w}_{t+1}=\frac{\alpha^{k}}{1+\alpha^{k}}\overline{v}_{t+1}+\frac{1}{1+\alpha^{k}}\overline{y}_{t+1}$.

For the linear regression problem in the interpolation setting, we
can write 
\begin{align*}
F(w) & =\frac{1}{2}(w-w^{\ast})^{T}H(w-w^{\ast})\\
 & =\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}
\end{align*}
 and similarly $F^{k}(w)=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}$,
so that 
\begin{align*}
g_{t,k} & =\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})\\
g_{t} & =\sum_{k=1}^{N}p_{k}\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})
\end{align*}

\subsection{Exponential Convergence when Global Loss is 0}

We now present the exponential convergence result using FedAvg with
MaSS updates. On each device, local data is stored and mini-batch
gradient descent with batch size $m^{k}$ is performed. We allow for
varying batch sizes across devices. 
\begin{thm}
(FedAvg with MaSS, Full Participation) Let the hyperparameters satisfy
the requirements of {[}Liu\&Belkin{]}. More precisely, let $\tilde{\kappa}_{m}:=\tilde{\kappa}/m^{k}+(m^{k}-1)/m^{k}$,
and let the hyper parameters satisfy 
\begin{align*}
\eta^{k}(m)=\frac{1}{L_{m}}, & \alpha^{k}(m)=\frac{1}{\sqrt{\kappa_{m}^{k}\tilde{\kappa}_{m}^{k}}},\delta^{k}(m)=\frac{\eta^{k}}{\alpha^{k}\tilde{\kappa}_{m}^{k}}
\end{align*}
Let $\mathcal{I}_{E}=\ell E$ where $\ell\in\mathbb{N}$ be the set
of communication rounds, then the updates 
\begin{align*}
v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}w_{t}^{k}-\delta^{k}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
 where $g_{t,k}$ are the mini-batch stochastic gradients achieves
the convergence
\begin{align*}
\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot(1-1/\sqrt{\kappa_{m}}\\
\end{align*}
\end{thm}
\begin{proof}
Note that at each communication round we update the $w_{t+1}^{k}$
parameters to be the average across devices while fixing $v_{t+1}^{k}$.
This automatically adjusts the $u_{t+1}^{k}$ parameter at each device
by the relation 
\begin{align*}
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
 valid for all $t\geq0$.

\textbf{For now assume $\delta^{k}=\delta$ and $\alpha^{k}=\alpha$
is equal across devices. Need to relax later. }

Theorems 2 and 3 of the Liu\&Belkin paper gives the bound 
\begin{align*}
\mathbb{E}\left[\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{E}^{k}-\eta^{k}g_{E,k}-w^{\ast}\|^{2}\right] & \leq(1-\alpha^{k})^{E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
for all $k$, where $E$ is the first communication round. Note that
$w_{E}^{k}=\overline{w}_{E}^{k}\neq u_{E}^{k}-\eta^{k}g_{t,k}$.

It follows from convexity that 
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{E}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\left[\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{E}^{k}-\eta^{k}g_{E,k}-w^{\ast}\|^{2}\right]\\
 & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
Since $w_{E}^{k}=\overline{w}_{E}$ for all devices, applying the
per-device result again starting at $t=E$ instead of $t=0$, for
each device we have the bound
\begin{align*}
\mathbb{E}\left[\|v_{2E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{2E}^{k}-\eta^{k}g_{2E,k}-w^{\ast}\|^{2}\right] & \leq(1-\alpha^{k})^{E}\mathbb{E}(\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{E}^{k}-w^{\ast}\|^{2})\\
 & =(1-\alpha^{k})^{E}\mathbb{E}(\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{E}-w^{\ast}\|^{2})
\end{align*}

Here we emphasize that $w_{E}^{k}$ results from broadcasting and
so is the same across all devices, while $v_{E}^{k}$ remains distinct
on each device (and is only auxiliary). Then by convexity and summing
the above inequalities across devices we have 
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{2E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{2E}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\left[\|v_{2E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{2E}^{k}-\eta^{k}g_{2E,k}-w^{\ast}\|^{2}\right]\\
 & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{E}\mathbb{E}(\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{E}-w^{\ast}\|^{2})\\
 & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{2E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
 and by induction we can show that 
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{\ell E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{\ell E}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{\ell E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
 and more generally
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{t}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{t}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{t}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
 In particular, this implies 
\begin{align*}
\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{t}
\end{align*}
 which is exponential convergence. 
\end{proof}

\end{document}